---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stat5555sor

<!-- badges: start -->
[![R-CMD-check](https://github.com/Gideonparry/stat5555sor/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Gideonparry/stat5555sor/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The stat5555sor package is designed with the intent to find the strength of record for the last 3 NCAA basketball seasons(21-23). The user can decide which season they would like to look at and through our algorithm, they can see which teams were actually the strongest during those seasons.

## Installation

You can install the development version of stat5555sor from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Gideonparry/stat5555sor")
```

## Date Frame Example

Before looking at the Data Frame, one should know that each row represents one game played during the selected season. In this case, it will be the 2022-2023 season.The variables are all explained below.

```{r, echo = FALSE, include = FALSE}
devtools::install_github("Gideonparry/stat5555sor", auth_token = "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", force = TRUE)
devtools::load_all()
```

```{r example}
test <- get_combined_data(2023)
head(test)
```

Variables:

Date: The day the game was played\
Away_team/Home_team: Name of teams playing in game\
Away_NET/Home_NET: Rank of that team among all 358 Division I basketball teams\
score_diff: Difference between Away_score and Home_score\
site: location the game was played\
neutral: refers to whether the game was played at the site of either team's home location or at a neutral site\
Away_record/Home_record: The record for each home and away team for the season selected for each team in that game\
